The dual frame of coherent states with random amplitudes

### 2013-10-05 dual.tex
writing

The illustrated approximation methods use Tychonoff regularised least squares, with different values of the regularisation parameter $λ$.  They assign expansion coefficients $d$ in order that
$$\pmatrix{|A〉\cr λI}d≈\pmatrix{|ψ〉\cr 0},$$
this equation being satisfied in a least-squares sense.  The solution minimises the residual $‖|ψ〉-|A〉d‖²+λ²‖d‖²$.  It is given by
$$d=〈A⁺|ψ〉=\pmatrix{|A〉\cr λI}⁺\pmatrix{1\cr 0}|ψ〉,$$
where $()⁺$ denotes the Moore-Penrose pseudoinverse of a matrix.  I.e., the vector $〈A⁺|$ is the leftmost column of the pseudoinverse matrix.

>>>>

The illustrated approximation methods use Tychonoff regularised least squares, with different values of the regularisation parameter $λ$.  They assign expansion coefficients $d$ in order that
$$\pmatrix{|A〉\cr λI}d≈\pmatrix{|ψ〉\cr 0},$$
this equation being satisfied in a least-squares sense.  The solution minimises the residual $‖|ψ〉-|A〉d‖²+λ²‖d‖²$.  It is given by
$$d=〈A⁺|ψ〉=\pmatrix{|A〉\cr λI}⁺\pmatrix{1\cr 0_R}|ψ〉,$$
where $()⁺$ denotes the Moore-Penrose pseudoinverse of a matrix, $I$ is the $R×R$ unit matrix, and $0_R$ is a $R×1$ zero vector.  I.e., the vector $〈A⁺|$ is the leftmost column of the pseudoinverse matrix.

Suppose $|A〉$ has the singular-value decomposition $|A〉=|U〉SV†$, so that
$$\pmatrix{|A〉\cr λI}v_r=\pmatrix{σ_r|u_r〉\cr λv_r}=\sqrt{σ_r²+λ²}w_r.$$
The orthonormality of the $w_r$ follows from that of the $|u_r〉$ and $v_r$, so we have a singular-value decomposition
$$\pmatrix{|A〉\cr λI}=WTV†,$$
where $τ_r=\sqrt{σ_r²+λ²}$.  Then 
$$〈A⁺|=\pmatrix{|A〉\cr λI}⁺\pmatrix{1\cr 0_R}=VT^{-1}W†\pmatrix{1\cr 0_R},$$
where the last two factors are the first column of $W†$,
$$W†\pmatrix{1\cr 0_R}=\pmatrix{{σ₁\over\sqrt{σ₁²+λ²}}〈u₁|\cr\vdots\cr 
	{σ_R\over\sqrt{σ_R²+λ²}}〈u_R|}.$$
It follows that the right singular vectors of $〈A⁺|$ are the bras $〈u_r|$, and the singular values are 
$$σ⁺_r={σ_r\over τ_r\sqrt{σ_r²+λ²}}={σ_r\over σ_r²+λ²}.$$
Tychonov regularisation divides the singular values of $|A〉$ into large ones and small ones, as compared to $λ$.  For large $σ_r$, the inverse is very close to the unregularised problem, with $σ⁺_r≈1/σ_r$.  For small $σ_r$, the component in that singular direction is nearly discarded by $|Α⁺〉$, with $σ⁺_r≈σ_r/λ²$.  Note that, in this limit, $|Α⁺〉=λ^{-2}|A〉$!  In between, there is a maximum at $σ_r=λ$, with $σ_r=1/2λ$, half the unregularised value.

This has dramatic consequences for the conditioning of the regularised problem, and how that varies with $λ$.  When $λ<σ_R$, regularisation has very little effect: the condition of $〈Α⁺|$ is the same as that of $|A〉$, which, for dense sets of coherent states, is usually dreadful.  As $λ$ increases, so that $σ₁<λ<σ_R$, the singular values lie on either side of the $σ_r=λ$ peak; the condition of $〈Α⁺|$ is determined by how far below this peak $σ₁$ and $σ_R$ fall, and the condition number has a sharp minimum when $σ⁺₁=σ⁺_R$.  When $λ$ dominates the singular values, and $|Α⁺〉=λ^{-2}|A〉$, it is again the case that their condition is the same.
